---
title: "Bikeshare vs Taxi: Pre-COVID and Post-COVID"
output: html_notebook
---

We want to look at the Capital Bikeshare program before and after COVID-19 compared to other for-hire vehicles's recovery after COVID-19. Here, we load our library files and data.

```{r}
library(tidyverse)
library(here)
library(janitor)
library(tmap)
library(sf)
library(ggplot2)
library(plotly)

dfApr20 = read_csv(here("data_raw",
                    "202004-capitalbikeshare-tripdata.zip"))

dfMay20 = read_csv(here("data_raw",
                    "202005-capitalbikeshare-tripdata.zip"))

dfJun20 = read_csv(here("data_raw",
                    "202006-capitalbikeshare-tripdata.zip"))

dfJul20 = read_csv(here("data_raw",
                    "202007-capitalbikeshare-tripdata.zip"))

dfAug20 = read_csv(here("data_raw",
                    "202008-capitalbikeshare-tripdata.zip"))

dfSep20 = read_csv(here("data_raw",
                    "202009-capitalbikeshare-tripdata.zip"))

dfOct20 = read_csv(here("data_raw",
                    "202010-capitalbikeshare-tripdata.zip"))

dfNov20 = read_csv(here("data_raw",
                    "202011-capitalbikeshare-tripdata.zip"))

dfDec20 = read_csv(here("data_raw",
                    "202012-capitalbikeshare-tripdata.zip"))

dfApr22 = read_csv(here("data_raw",
                    "202204-capitalbikeshare-tripdata.zip"))

dfMay22 = read_csv(here("data_raw",
                    "202205-capitalbikeshare-tripdata.zip"))

dfJun22 = read_csv(here("data_raw",
                    "202206-capitalbikeshare-tripdata.zip"))

dfJul22 = read_csv(here("data_raw",
                    "202207-capitalbikeshare-tripdata.zip"))

dfAug22 = read_csv(here("data_raw",
                    "202208-capitalbikeshare-tripdata.zip"))

dfSep22 = read_csv(here("data_raw",
                    "202209-capitalbikeshare-tripdata.zip"))

dfOct22 = read_csv(here("data_raw",
                    "202210-capitalbikeshare-tripdata.zip"))

dfNov22 = read_csv(here("data_raw",
                    "202211-capitalbikeshare-tripdata.zip"))

dfDec22 = read_csv(here("data_raw",
                    "202212-capitalbikeshare-tripdata.zip"))
```

Clean and make the data about cumulative riders
```{r}
dfAll20 = bind_rows(dfApr20, dfMay20, dfJun20, dfJul20, dfAug20, dfSep20, dfOct20, dfNov20, dfDec20)

dfAll22 = bind_rows(dfApr22, dfMay22, dfJun22, dfJul22, dfAug22, dfSep22, dfOct22, dfNov22, dfDec22)

dfAll20Start = dfAll20 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll20End = dfAll20 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll20Riders = bind_rows(dfAll20Start,
                dfAll20End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") *1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)

dfAll22Start = dfAll22 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll22End = dfAll22 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll22Riders = bind_rows(dfAll22Start,
                dfAll22End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") *1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

Plot of riders vs time:
```{r}
dfAll20Riders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
dfAll22Riders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```


Now we load the taxi data:

```{r}
dfApr20 = read_csv(here("data_raw",
                    "taxi_2020_04.zip"))

dfMay20 = read_csv(here("data_raw",
                    "taxi_2020_05.zip"))

dfJun20 = read_csv(here("data_raw",
                    "taxi_2020_06.zip"))

dfJul20 = read_csv(here("data_raw",
                    "taxi_2020_07.zip"))

dfAug20 = read_csv(here("data_raw",
                    "taxi_2020_08.zip"))

dfSep20 = read_csv(here("data_raw",
                    "taxi_2020_09.zip"))

dfOct20 = read_csv(here("data_raw",
                    "taxi_2020_10.zip"))

dfNov20 = read_csv(here("data_raw",
                    "taxi_2020_11.zip"))

dfDec20 = read_csv(here("data_raw",
                    "taxi_2020_12.zip"))
dfApr22 = read_csv(here("data_raw",
                    "taxi_202204.zip"))

dfMay22 = read_csv(here("data_raw",
                   "taxi_202205.zip"))

dfJun22 = read_csv(here("data_raw",
                    "taxi_202206.zip"))

dfJul22 = read_csv(here("data_raw",
                    "taxi_202207.zip"))

dfAug22 = read_csv(here("data_raw",
                    "taxi_202208.zip"))

dfSep22 = read_csv(here("data_raw",
                   "taxi_202209.zip"))

dfOct22 = read_csv(here("data_raw",
                   "taxi_202210.zip"))

dfNov22 = read_csv(here("data_raw",
                    "taxi_202211.zip"))

dfDec22 = read_csv(here("data_raw",
                   "taxi_202211.zip"))

dfAll20 = bind_rows(dfApr20, dfMay20, dfJun20, dfJul20, dfAug20, dfSep20, dfOct20, dfNov20, dfDec20)

dfAll22 = bind_rows(dfApr22, dfMay22, dfJun22, dfJul22, dfAug22, dfSep22, dfOct22, dfNov22, dfDec22)
```

Plot of total taxi users versus time.
```{r}
dfAll20Users = dfAll20 |>
  mutate(users_delta = (MILEAGE != 0)*1) |>
  mutate(user = cumsum(users_delta))
dfAll20Users = dfAll20Users |> 
  rename(r = user,
    t = ORIGINDATETIME_TR)

dfAll22Users = dfAll22 |>
  mutate(users_delta = (MILEAGE != 0)*1) |>
  mutate(user = cumsum(users_delta))
dfAll22Users = dfAll22Users |> 
  rename(r = user,
    t = ORIGINDATETIME_TR)
```
Plot of taxi users vs time:
```{r}
dfAll20Users %>% 
  ggplot(aes(t, r)) +
  geom_line()
dfAll22Users %>% 
  ggplot(aes(t, r)) +
  geom_line()
```
Interactive plot of riders vs time:
```{r}
Fig20<- plot_ly(data = dfAll20Riders, x = ~t) |> add_trace(y = ~riders, name = 'Ridership', mode = 'lines')
Fig20

Fig22<- plot_ly(data = dfAll22Riders, x = ~t) |> add_trace(y = ~riders, name = 'Ridership', mode = 'lines')
Fig22
```

Interactive plot with taxi data:
```{r}
TFig20<- plot_ly(data = dfAll20Users, x = ~t) |> add_trace(y = ~r, name = 'Taxi ridership', mode = 'lines')
TFig20

TFig22<- plot_ly(data = dfAll22Users, x = ~t) |> add_trace(y = ~r, name = 'Taxi ridership', mode = 'lines')
TFig22
```