---
title: "2022 Bikeshare Data Analysis Project"
author: "Group Project"
date: "2023-12-08"
output: html_notebook
---

Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(tmap)
library(sf)
library(ggplot2)
library(plotly)
```

We will begin creating data frames for every month's Bikeshare .zip file for year 2020. We are excluding January, Feb., and March because the data is broken from the source.

```{r}
dfApr20 = read_csv(here("data_raw",
                    "202004-capitalbikeshare-tripdata.zip"))

dfMay20 = read_csv(here("data_raw",
                    "202005-capitalbikeshare-tripdata.zip"))

dfJun20 = read_csv(here("data_raw",
                    "202006-capitalbikeshare-tripdata.zip"))

dfJul20 = read_csv(here("data_raw",
                    "202007-capitalbikeshare-tripdata.zip"))

dfAug20 = read_csv(here("data_raw",
                    "202008-capitalbikeshare-tripdata.zip"))

dfSep20 = read_csv(here("data_raw",
                    "202009-capitalbikeshare-tripdata.zip"))

dfOct20 = read_csv(here("data_raw",
                    "202010-capitalbikeshare-tripdata.zip"))

dfNov20 = read_csv(here("data_raw",
                    "202011-capitalbikeshare-tripdata.zip"))

dfDec20 = read_csv(here("data_raw",
                    "202012-capitalbikeshare-tripdata.zip"))
```

We will then create data frames for every month's Bikeshare .zip file for year 2022. We are excluding Jan., Feb., and March 2022, since we are also excluding them from 2020.

```{r}
dfApr22 = read_csv(here("data_raw",
                    "202204-capitalbikeshare-tripdata.zip"))

dfMay22 = read_csv(here("data_raw",
                    "202205-capitalbikeshare-tripdata.zip"))

dfJun22 = read_csv(here("data_raw",
                    "202206-capitalbikeshare-tripdata.zip"))

dfJul22 = read_csv(here("data_raw",
                    "202207-capitalbikeshare-tripdata.zip"))

dfAug22 = read_csv(here("data_raw",
                    "202208-capitalbikeshare-tripdata.zip"))

dfSep22 = read_csv(here("data_raw",
                    "202209-capitalbikeshare-tripdata.zip"))

dfOct22 = read_csv(here("data_raw",
                    "202210-capitalbikeshare-tripdata.zip"))

dfNov22 = read_csv(here("data_raw",
                    "202211-capitalbikeshare-tripdata.zip"))

dfDec22 = read_csv(here("data_raw",
                    "202212-capitalbikeshare-tripdata.zip"))

```

Now, we will combine all of the data frames into yearly data frames for years 2020 and 2022.

```{r}
dfAll20 = bind_rows(dfApr20, dfMay20, dfJun20, dfJul20, dfAug20, dfSep20, dfOct20, dfNov20, dfDec20)
```

```{r}
dfAll22 = bind_rows(dfApr22, dfMay22, dfJun22, dfJul22, dfAug22, dfSep22, dfOct22, dfNov22, dfDec22)
```

Now, we will create a visualization for each yearly data frame, years 2020 and 2022.

```{r}
dfAll20Start = dfAll20 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll20End = dfAll20 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll20Riders = bind_rows(dfAll20Start,
                dfAll20End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") *1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)

dfAll22Start = dfAll22 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll22End = dfAll22 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll22Riders = bind_rows(dfAll22Start,
                dfAll22End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") *1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

```{r}
Fig20<- plot_ly(data = dfAll20Riders, x = ~t) |> add_trace(y = ~riders, name = 'Ridership', mode = 'lines')
#Fig20

Fig22<- plot_ly(data = dfAll22Riders, x = ~t) |> add_trace(y = ~riders, name = 'Ridership', mode = 'lines')
#Fig22
```

Derivative plot: How much riders change over time

```{r}
original20 <- data.frame(X = dfAll20Riders$t, Y = dfAll20Riders$riders, type = "Original")
original20<- original20[-1,]
diffx <- diff(dfAll20Riders$t)
diffx<- as.numeric(diffx)
diffy <- diff(dfAll20Riders$riders)
ridersdiff = diffy/diffx

do20<- original20 |> mutate(ridersDiff = ridersdiff)

ggplot(do20, aes(x=X)) + geom_point(aes(y= Y), color = "black") + geom_point(aes(y=ridersdiff), color = "darkred") 
```
