---
title: "2022 Bikeshare Data Analysis Project"
author: "Group Project"
date: "2023-12-08"
output: html_notebook
---

Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(tmap)
library(sf)
library(ggplot2)
library(plotly)
```

We will begin creating data frames for every month's Bikeshare .zip file for year 2020.

```{r}
dfJan20 = read_csv(here("data_raw",
                    "202001-capitalbikeshare-tripdata.zip"))

dfFeb20 = read_csv(here("data_raw",
                    "202002-capitalbikeshare-tripdata.zip"))

dfMar20 = read_csv(here("data_raw",
                    "202003-capitalbikeshare-tripdata.zip"))

dfApr20 = read_csv(here("data_raw",
                    "202004-capitalbikeshare-tripdata.zip"))

dfMay20 = read_csv(here("data_raw",
                    "202005-capitalbikeshare-tripdata.zip"))

dfJun20 = read_csv(here("data_raw",
                    "202006-capitalbikeshare-tripdata.zip"))

dfJul20 = read_csv(here("data_raw",
                    "202007-capitalbikeshare-tripdata.zip"))

dfAug20 = read_csv(here("data_raw",
                    "202008-capitalbikeshare-tripdata.zip"))

dfSep20 = read_csv(here("data_raw",
                    "202009-capitalbikeshare-tripdata.zip"))

dfOct20 = read_csv(here("data_raw",
                    "202010-capitalbikeshare-tripdata.zip"))

dfNov20 = read_csv(here("data_raw",
                    "202011-capitalbikeshare-tripdata.zip"))

dfDec20 = read_csv(here("data_raw",
                    "202012-capitalbikeshare-tripdata.zip"))
```

We will then create data frames for every month's Bikeshare .zip file for year 2022.

```{r}
dfJan22 = read_csv(here("data_raw",
                    "202201-capitalbikeshare-tripdata.zip"))

dfFeb22 = read_csv(here("data_raw",
                    "202202-capitalbikeshare-tripdata.zip"))

dfMar22 = read_csv(here("data_raw",
                    "202203-capitalbikeshare-tripdata.zip"))

dfApr22 = read_csv(here("data_raw",
                    "202204-capitalbikeshare-tripdata.zip"))

dfMay22 = read_csv(here("data_raw",
                    "202205-capitalbikeshare-tripdata.zip"))

dfJun22 = read_csv(here("data_raw",
                    "202206-capitalbikeshare-tripdata.zip"))

dfJul22 = read_csv(here("data_raw",
                    "202207-capitalbikeshare-tripdata.zip"))

dfAug22 = read_csv(here("data_raw",
                    "202208-capitalbikeshare-tripdata.zip"))

dfSep22 = read_csv(here("data_raw",
                    "202209-capitalbikeshare-tripdata.zip"))

dfOct22 = read_csv(here("data_raw",
                    "202210-capitalbikeshare-tripdata.zip"))

dfNov22 = read_csv(here("data_raw",
                    "202211-capitalbikeshare-tripdata.zip"))

dfDec22 = read_csv(here("data_raw",
                    "202212-capitalbikeshare-tripdata.zip"))

```

Now, we will combine all of the data frames into yearly data frames for years 2020 and 2022.
```{r}
dfJan20$`End station number` = as.numeric(dfJan20$`End station number`)
dfAll20 = bind_rows(dfJan20, dfFeb20, dfMar20, dfApr20, dfMay20, dfJun20, dfJul20, dfAug20, dfSep20, dfOct20, dfNov20, dfDec20)
```

```{r}
dfAll22 = bind_rows(dfJan22, dfFeb22, dfMar22, dfApr22, dfMay22, dfJun22, dfJul22, dfAug22, dfSep22, dfOct22, dfNov22, dfDec22)
```

Now, we will create a visualization for each yearly data frame, years 2020 and 2022.

```{r}
dfAll20Start = dfAll20 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll20End = dfAll20 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll20Riders = bind_rows(dfAll20Start,
                dfAll20End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

THIS GRAPH IS NOT WORKING:

```{r}
dfAll20Riders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```

```{r}
dfAll22Start = dfAll22 %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfAll22End = dfAll22 %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfAll22Riders = bind_rows(dfAll22Start,
                dfAll22End) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

```{r}
dfAll22Riders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```

Month by Month, 2020 (MONTHS unfinished/unedited)

```{r}
dfJanStart = dfJan %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfJanEnd = dfJan %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfJanRiders = bind_rows(dfJanStart,
                dfJanEnd) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

```{r}
dfJanRiders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```

```{r}
dfFebStart = dfFeb %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfFebEnd = dfFeb %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfFebRiders = bind_rows(dfFebStart,
                dfFebEnd) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

```{r}
dfFebRiders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```


```{r}
dfDecStart = dfDec %>% 
  select(rideable_type,
         member_casual,
         contains("start"),
         ride_id) %>% 
  mutate(start_stop = "start") %>%
  rename(t = started_at,
         station_name = start_station_name,
         station_id = start_station_id,
         lat = start_lat,
         lng = start_lng)

dfDecEnd = dfDec %>% 
  select(ride_id,
         rideable_type,
         member_casual,
         contains("end")) %>%
  mutate(start_stop = "stop") %>%
  rename(t = ended_at,
         station_name = end_station_name,
         station_id = end_station_id,
         lat = end_lat,
         lng = end_lng)
 
dfDecRiders = bind_rows(dfDecStart,
                dfDecEnd) %>%
  arrange(t) %>%
  mutate(rider_delta = (start_stop == "start") * 2 - 1) %>% # change in ridership 
  mutate(riders = cumsum(rider_delta)) %>%
  relocate(riders, .after = t)
```

```{r}
dfDecRiders %>% 
  ggplot(aes(t, riders)) +
  geom_line()
```

