---
title: "2022 Bikeshare Data Analysis Project, Taxi"
author: "Group Project"
date: "2023-12-08"
output: html_notebook
---
Adjusted
Packages

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(tmap)
library(sf)
library(plotly)
```

We will begin creating data frames for every month's Bikeshare .zip file for year 2020. We are excluding January, Feb., and March because the data is broken from the source.

```{r}
dfApr20 = read_csv(here("data_raw",
                    "taxi_2020_04.zip"))

dfMay20 = read_csv(here("data_raw",
                    "taxi_2020_05.zip"))

dfJun20 = read_csv(here("data_raw",
                    "taxi_2020_06.zip"))

dfJul20 = read_csv(here("data_raw",
                    "taxi_2020_07.zip"))

dfAug20 = read_csv(here("data_raw",
                    "taxi_2020_08.zip"))

dfSep20 = read_csv(here("data_raw",
                    "taxi_2020_09.zip"))

dfOct20 = read_csv(here("data_raw",
                    "taxi_2020_10.zip"))

dfNov20 = read_csv(here("data_raw",
                    "taxi_2020_11.zip"))

dfDec20 = read_csv(here("data_raw",
                    "taxi_2020_12.zip"))
```

We will then create data frames for every month's Taxi .zip file for year 2022. We are excluding Jan., Feb., and March 2022, since we are also excluding them from 2020.

```{r}
dfApr22 = read_csv(here("data_raw",
                    "taxi_202204.zip"))

dfMay22 = read_csv(here("data_raw",
                   "taxi_202205.zip"))

dfJun22 = read_csv(here("data_raw",
                    "taxi_202206.zip"))

dfJul22 = read_csv(here("data_raw",
                    "taxi_202207.zip"))

dfAug22 = read_csv(here("data_raw",
                    "taxi_202208.zip"))

dfSep22 = read_csv(here("data_raw",
                   "taxi_202209.zip"))

dfOct22 = read_csv(here("data_raw",
                   "taxi_202210.zip"))

dfNov22 = read_csv(here("data_raw",
                    "taxi_202211.zip"))

dfDec22 = read_csv(here("data_raw",
                   "taxi_202211.zip"))

```

Now, we will combine all of the data frames into yearly data frames for years 2020 and 2022.

```{r}
dfAll20 = bind_rows(dfApr20, dfMay20, dfJun20, dfJul20, dfAug20, dfSep20, dfOct20, dfNov20, dfDec20)
```

```{r}
dfAll22 = bind_rows(dfApr22, dfMay22, dfJun22, dfJul22, dfAug22, dfSep22, dfOct22, dfNov22, dfDec22)
```

Now, we will create a visualization for each yearly data frame, years 2020 and 2022.

create user_data column
2020 graph
```{r}
dfAll20Users = dfAll20 |>
  mutate(users_delta = (MILEAGE != 0)*1) |>
  mutate(user = cumsum(users_delta))
dfAll20Users = dfAll20Users |> 
  rename(r = user,
    t = ORIGINDATETIME_TR)
dfAll20Users %>% 
  ggplot(aes(t, r)) +
  geom_line()
```
2022 graph
```{r}
dfAll22Users = dfAll22 |>
  mutate(users_delta = (MILEAGE != 0)*1) |>
  mutate(user = cumsum(users_delta))
dfAll22Users = dfAll22Users |> 
  rename(r = user,
    t = ORIGINDATETIME_TR)
dfAll22Users %>% 
  ggplot(aes(t, r)) +
  geom_line()
```
